---
title: "Cookie: representative samples selection from large population by using k-medoids clustering "
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load package
```{r results='hide', message=FALSE, warning=FALSE}
library(Cookie)
library(cowplot)
library(ggplot2)
```

# Load data
User can use read.csv() to load data into R. Columns should be factors, and each sample should have an unique name or ID (in this example is cell barcode). 
The missing values should be indicated by "NA" in CSV file. Users can arrange and edit their data in Excel and then import into R. 
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
data <- read.csv(file = system.file("extdata", "data.csv", package = "Cookie"), header = TRUE, row.names = 1)
```

# Create object
Creat data object and determine factor types
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
cookie <- createCookieObject(data=data,factor.type = c("char","char","char","char","num"))
```

# Normalize data (for numerical factors)
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
cookie <- normalization(cookie)
```

# Calculate pair-wise distances
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
cookie <- distCalculation(cookie)
```

# Generate non-linear dim reduction
Users also can use UMAP (https://github.com/tkonopka/umap) to generate non-linear reduction. Two mehoeds is avaiolable: "naive" and "umap-learn". To use "umap-learn", users may need to install python package "umap-learn" (https://pypi.org/project/umap-learn/). 
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
cookie <- reductionTSNE(cookie)
#cookie <- reductionUMAP(cookie)
```

# See how factors distributed
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotReduction(cookie, reduction = "tsne", group.factor = "Subject") 
p2 <- plotReduction(cookie, reduction = "tsne", group.factor = "Isotype") 
plot_grid(p1,p2,labels = c("Subject","Isotype"), ncol = 2)
```

# Test coverages for different sample size to determine an appropriate sample size
we run two test 1) with prime factor 2) without prime factor
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
cookie <- sampleSizeTest(cookie, prime.factor = "Subject",size.range = c(3,5,7,9,11,13,15),name = "run1")

# When selecting samples from entire large population, users can use fast.mode = 3, 4, or 5 to enable FastPAM algorithm.  
cookie <- sampleSizeTest(cookie, size.range = c(60,70,80,90,100,110,120),name = "run2", fast.mode = 5)
```
### 1) with prime factor, coverage rate of each sample size
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotSizeTest(cookie, test.name = "run1", chart.type = "line", coverage.type = "pct")
p2 <- plotSizeTest(cookie, test.name = "run1", chart.type = "radar", coverage.type = "pct")

plot_grid(p1,p2,ncol = 2, rel_widths = c(1,1.5))
```
### 2) without prime factor, coverage rate of each sample size
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotSizeTest(cookie, test.name = "run2", chart.type = "line", coverage.type = "pct")
p2 <- plotSizeTest(cookie, test.name = "run2", chart.type = "radar", coverage.type = "pct")

plot_grid(p1,p2,ncol = 2, rel_widths = c(1,1.5))
```
# After determine an appropriate sample size, users can select samples from original population
we run two sampling 1) with prime factor 2) without prime factor
```{r}
cookie <- sampling(cookie, prime.factor = "Subject",important.factor = c("Isotype"),sample.size = 5,name = "run1")

# When selecting samples from entire large population, users can use fast.mode = 3, 4, or 5 to enable FastPAM algorithm.(https://stat.ethz.ch/R-manual/R-patched/library/cluster/html/pam.html)
cookie <- sampling(cookie, important.factor = c("Isotype","Subject"),sample.size = 90, name = "run2", fast.mode = 5)
```
Coverage on each factor of current sampling 1) with prime factor
Because all samples were selected evenly from each subject, the STD of selected samples on subject is 0. The CC value is not available. We set CC to 0. 
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotSampling(cookie,name = "run1", coverage.type = "pct") + theme(legend.position = "none")
p2 <- plotSampling(cookie,name = "run1", coverage.type = "cc") + theme(legend.position = "none")

plot_grid(p1,p2,labels = c("Coverage rates","CC values between original and selection"), ncol = 2)
```

### Coverage on each factor of current sampling 2) without prime factor
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotSampling(cookie,name = "run2", coverage.type = "pct") + theme(legend.position = "none")
p2 <- plotSampling(cookie,name = "run2", coverage.type = "cc") + theme(legend.position = "none")

plot_grid(p1,p2,labels = c("Coverage rates","CC values between original and selection"), ncol = 2)
```

### Selected samples loacted on 2D map
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotReduction(cookie, reduction = "tsne", group.factor = "Subject", sampling = "run1") 
p2 <- plotReduction(cookie, reduction = "tsne", group.factor = "Isotype", sampling = "run1") 
plot_grid(p1,p2,labels = c("Subject","Isotype"), ncol = 2)
```

Distribution of original population and current sampling
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
p1 <- plotDistribution(cookie, sampling = "original",factor = "Subject") + labs(title = "")
p2 <- plotDistribution(cookie, sampling = "original",factor = "Seuret_Cluster") + labs(title = "")
p3 <- plotDistribution(cookie, sampling = "original",factor = "Isotype") + labs(title = "")
p4 <- plotDistribution(cookie, sampling = "original",factor = "V.locus") + labs(title = "")
p5 <- plotDistribution(cookie, sampling = "original",factor = "CDR3.length") + labs(title = "")

p6 <- plotDistribution(cookie, sampling = "run2",factor = "Subject") + labs(title = "")
p7 <- plotDistribution(cookie, sampling = "run2",factor = "Seuret_Cluster") + labs(title = "")
p8 <- plotDistribution(cookie, sampling = "run2",factor = "Isotype") + labs(title = "")
p9 <- plotDistribution(cookie, sampling = "run2",factor = "V.locus") + labs(title = "")
p10 <- plotDistribution(cookie, sampling = "run2",factor = "CDR3.length") + labs(title = "")

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10, ncol = 5, labels = c("Original-Subject","Original-Cluster","Original-Isotype","Original-V.locus","Original-CDR3.length","Sampling-Subject","Sampling-Cluster","Sampling-Isotype","Sampling-V.locus","Sampling-CDR3.length"))
```

